\chapter*{Preface}\label{chapter:preface}

This thesis was conducted as part of the requirements for obtaining my Master’s degree. It represents the culmination of my academic journey, combining both the theoretical knowledge I have acquired over the years from the professors who taught me as well as the practical experience I gained from my time at the student team E.S.A.I.V Serpentine, my professional engagement at HWKY, and my research at Avular.

\paragraph*{}The latter provided me with invaluable guidance and support throughout the course of this thesis. I am especially grateful to Bart Keulen, whose expertise and insights were instrumental in shaping my research. Our many fruitful discussions significantly enhanced the quality of this work. I also want to express my sincere thanks for his meticulous review of my work, which greatly contributed to its clarity and overall quality.

\paragraph*{}I wish to extend my heartfelt thanks to Jakub Tomczak for generously sharing his extensive knowledge of the subject. His expertise was invaluable, particularly in resolving the technical challenges I encountered throughout my research. Jakub's willingness to assist and provide guidance played a crucial role in overcoming these obstacles. Additionally, I am grateful to Eindhoven University of Technology for supplying the essential computational resources necessary for conducting all experiments. Their support in providing the infrastructure and tools required was instrumental in the successful completion of this thesis.

\paragraph*{}My peers from the study association also deserve special mention. The countless discussions we had and the thorough reviews they provided helped improve the readability of my thesis and ensured that errors were caught early. Their contributions made a significant difference, and for that, I am truly grateful.

\paragraph*{}I would also like to thank my roommates, who took over the cooking duties, allowing me to focus fully on my research. Their support and understanding made this journey smoother and more manageable.

\paragraph*{}Lastly, I wish to express my deepest and most heartfelt gratitude to my parents. Their steadfast support, encouragement, and unwavering belief in my abilities have been foundational to my personal and academic growth. Throughout my life, their constant presence and faith in my potential have provided me with the strength and motivation needed to pursue and achieve my goals. This accomplishment is a testament to their enduring support and sacrifices. Without their unwavering commitment and encouragement, this achievement would not have been possible.

\paragraph*{}As I present this thesis, I do so with a deep sense of accomplishment and gratitude to all those who have supported me along the way.